\documentclass[11pt, a4paper]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage{amsmath} 
\usepackage{algorithm}
\usepackage{algpseudocode}

\title{Project Title}
\author{Sam \and Charlie \and Anusha \and Mukund}
\date{}

\begin{document}

\maketitle

\section*{Project Goal and Scope}

The primary objective is to develop a photo-realistic 3D graphics rendering system in C++ utilizing advanced Path Tracing Techniques. 

The specific target for the rendering competition is to accurately recreate an image of a chess scene \cite{dartmouth_rendering}. Achieving this level of visual fidelity requires a synthesis of custom geometric modeling, designing material properties, and a Path Tracer that handles shadows, and glass surfaces.


\section*{Accomplishments}
\subsection*{Bidirectional Path Tracing }
We upgraded the core rendering engine to utilize Bidirectional Path Tracing \cite{Vlnas2018}. In traditional Monte Carlo path tracing, rays are exclusively cast from the camera into the scene. Our implementation improves upon this by simultaneously generating sub-paths from both the camera (eye paths) and the emissive light sources (light paths). The length of both sub-paths is determined stochastically, utilizing Russian Roulette for unbiased termination.


\section*{Technical Details}

To evaluate the final pixel color, the integrator attempts to connect the vertices of the camera sub-path with the vertices of the light sub-path. For each attempted connection, a deterministic visibility check (shadow ray) is cast between the two endpoints. If an object occludes the line of sight between the vertices, no color contribution is added for that specific connection. If the path is unoccluded, the radiance is evaluated and accumulated similarly to standard Monte Carlo integration. 

\subsection*{Path Tracing vs. Bidirectional Path Tracing}

In standard unidirectional Path Tracing, the rendering equation is integrated exclusively by gathering light. Rays are cast from the camera into the scene, scattering at surfaces according to the material's BRDF until they randomly intersect an emissive light source or terminate via Russian Roulette. The final color at a surface is evaluated by accumulating the emitted radiance from the final hit point, attenuated by the material reflectance and geometric terms along the reverse path. However, this Monte Carlo integration method often struggles with complex indirect lighting, leading to high variance (noise) because camera paths rarely hit the light source purely by chance.

The rendering equation used to describe traditional path tracing 
\begin{equation}\label{eq:rendering}
    L_o(p, \omega_o) = L_e(p, \omega_o) + \int_{\Omega} f_r(p, \omega_i, \omega_o) L_i(p, \omega_i) (\omega_i \cdot n) \, d\omega_i
\end{equation}
The previous equation is evaluated in Recursively. A ray can be traced from the camera eye into the nearest hit point $x$, then sample a new direction $\omega_i$ above $x$ and repeat these steps until  some kind of termination condition, such as Russian roulette,

The rendering equation (\ref{eq:rendering}) can be transformed into a path integral formulation where instead of integrating over all directions of incoming light over the point $p$ , we integrate over points on the surfaces of objects. 

\begin{equation}\label{eq:rendering_path_integral)}
    L_p = \int_{D} f_j(\vec{p})  \, dD
\end{equation}
where $L_p$ is the radiance that flows through a pixel , $D$ is all possible light paths and $\vec{p}$ is all possible light paths.

This can be approximated using monte-carlo approximation as follows 

\begin{equation}\label{eq:rendering_path_integral_monte_carlo)}
    L_p = \frac{1}{N} \sum_{i=0}^{N}\frac{f_j(\vec{p_i})}{p(\vec{p_i})}
\end{equation}

where $p(\vec{p_i}) $ is the pdf of sampling path $\vec{p_i}$. The conversion into a pdf with respect to surface area $p(x)$ is written as:

\begin{equation}\label{eq:conversion_pdf)}
    p(x) = p(\omega) \left(\frac{N_{x'} \cdot \vec{x'x''}}{\|x'' - x'\|}\right)
\end{equation}

where $p{(\omega)}$ is the solid angle. Moreover , We know that 
\begin{equation}\label{eq:measurement)}
    L_r(x,\omega_0) = L_{direct} + L_{indirect}
\end{equation}
 
(\ref{eq:measurement)}) can be written with conversion to surface area as 

\begin{equation}
    L_p = \frac{1}{N}\sum_{i=0}^{N}L_{p_i}
\end{equation}

where $L_{p_i}$ is the radiance of a single sample which is calculated by 
\begin{equation}
    L_{p_i} = \sum_{t=0}^{N_E}C_t
\end{equation}
Where $C_t$ is the contribution of a single path of length $t$ and $N_E$ is the maximum path length. Applying monte carlo integration on the contribution gives us
\begin{equation}\label{eq:pathtracer}
    C_t = \frac{L_e(y_0,\overrightarrow{y_0x_t})}{p(y_0)}f_r(x_t,\overrightarrow{x_ty_0},\overrightarrow{x_tx_{t-1}})G(x_t \leftrightarrow y_0) \prod_{i=1}^{t-1}\frac{f_r(x_i,\overrightarrow{x_ix_{i+1}},\overrightarrow{x_ix_{i-1}}) | N_{x_i} \cdot \overrightarrow{x_ix_{i+1}}}{p(\overrightarrow{x_ix_{i+1}})}
\end{equation}

The first part of the equation corresponds to the direct lighting from $y_0$ onto  $x_t$,where $L_e(y_0,\overrightarrow{y_0x_t})$, is the amount of radiance from point $y_0$ going in the direction towards $x_t$. Then it is multiplied with the geometric term, and also with the BRDF. The rest of the equation belongs to indirect lighting. It is computed for each point on the path as the product of BRDF and the dot product, divided by the PDF relative to BRDF. 

The geometric coupling term can be defined as 

\begin{equation}
    G(x \leftrightarrow x') = V(x \leftrightarrow x') \frac{|N_x \cdot \overrightarrow{x'x}||N_{x'} \cdot \overrightarrow{xx'}|}{\|x-x'\|^2}
\end{equation}

where $V(x \leftrightarrow x')$ is a visibility function, which is equal to 1 if a ray acan be shot directly from point $x$ to $x'$ without hitting anything else, otherwise it is 0

The light tracing which represents a light path can be written using the path integral framework (\ref{eq:rendering_path_integral)}) as follows

\begin{equation}
    L_p = \sum_{s=0}^{N_L}C_s \frac {\|P_{pixel} - P_{eye} \|^2}{(\cos{\theta})^2A_{pixel}}
\end{equation}

where the sum represents path, from point $y_0$ to the maximum path length $N_L$,$C_s$ is the single path contribution and the rest of the equation is the conversion from flux to radiance. Contribution $C_s$ can be computed as

\begin{equation}\label{eq:lighttracing}
    C_s = \frac{L_e(y_0,\overrightarrow{y_0y_1})}{p(y_0,\overrightarrow{y_0y_1})}f_r(y_s,\overrightarrow{y_sy_{s+1}})G(y_s \leftrightarrow x_0) \frac{W_e(x_0,\overrightarrow{y_tx_0})}{p(x_0)} \left( \prod_{i=1}^{s-1}\frac{f_r(y_i,\overrightarrow{y_iy_{i+1}},\overrightarrow{y_iy_{i-1}})|N_{y_i} \cdot \overrightarrow{y_iy_{i+1}}}{p(\overrightarrow{y_iy_{i+1}})}\right)
\end{equation}

% FIX 1: Added $ around \overrightarrow{y_0y_1} in this paragraph
The first part is the emmitted light $L_e(y_0,\overrightarrow{y_0y_1})$ from the light source at point $y_0$. The PDF $p(y_0,\overrightarrow{y_0y_1})$ is probability of selecting a ray in the direction $\overrightarrow{y_0y_1}$. Considering a light that has uniform distribution of the emitted light, then probability density function is:
\begin{equation}
    p(y_0,\overrightarrow{y_0y_1}) = \frac{1}{A_{light}2\pi}
\end{equation}
$y_0$ can be sampled over the light area $A_{light}$ in all directions over the hemisphere above the point $y_0$. $W_e(x_0,\overrightarrow{y_tx_0})$ is the potential of the light ray going straight from the light source to the pixel, which is equal to 1, when the ray is in front of the camera, and when using a pinhole camera $p(x_0)$ is equal to 1 as well, becasue only $x_0$ can be chosen as a point.

% FIX 2: Added $ around \overrightarrow{y_iy_{i+1}} in this paragraph
The next part $f_r(y_s,\overrightarrow{y_sy_{s+1}},\overrightarrow{y_sy_{s+1}}) G(y_s \leftrightarrow x_0)$ represents a direct ray from $y_s$ to the eye $x_0$. The third part  of the equation is the product of: the BRDF,multiplied with the dot product between normal and direction $\overrightarrow{y_iy_{i+1}}$ and normalized by the PDF.

We have defined both the strategies of ray evaluation, and now Bidirectional path tracing can be defined as 

\begin{equation}
    L_p = \sum_{s=0}^{N_L}\sum_{t=0}^{N_E} w_{s,t} \cdot C_{s,t}
\end{equation}

where $C_{s,t}$ is the unweighted contribution of a path with $s$ vertices on the light path and $t$ vertices on the eye path. A visibility check between $s^{th}$ point on the eye path and $t^{th}$ point on the ligth path is a part of the geometric term in contribution function. And finally, $w_{s,t}$ is the weight function.

As it can be seen, evaluation of the $C_{s,t}$ is independent on the eye and light path; therefore, there are four important cases:

\begin{itemize}
    \item{$s=0,t=0$ light is directly visible from the eye, the evaluation can be done with direct path between eye the point $x_0$ and $y_0$}
    \item{$s>0,t=0$ in this case use light tracing defined in (\ref{eq:lighttracing}) }
    \item{$s=0,t>0$ corresponds to the classic path tracing case defined in (\ref{eq:pathtracer}) }
    \item{$s>0,t>0$ the final case is the evaluation of radiance from the eye path $t$ connected with $s$ vertices of the light path, reaching the pixel, defined as follows}
\end{itemize}

    \begin{equation}
        \begin{split} 
            C_{s,t} &= \frac{L_e(y_0,\overrightarrow{y_0y_1})}{p(y_0,\overrightarrow{y_0y_1})}G(y_s \leftrightarrow x_t)f_r(x_t,\overrightarrow{x_ty_s},\overrightarrow{x_tx_{t-1}})f_r(y_s,\overrightarrow{y_sy_{s+1}},\overrightarrow{y_sx_t}) \\
            &\quad \times \left(\prod_{i=1}^{t-1}\frac{f_r(x_i,\overrightarrow{x_ix_{i+1}},\overrightarrow{x_ix_{i-1}})|N_{x_i} \cdot \overrightarrow{x_ix_{i+1}}|}{p(\overrightarrow{x_ix_{i+1}})} \right) \\
            &\quad \times \left(\prod_{i=1}^{s-1}\frac{f_r(y_i,\overrightarrow{y_iy_{i+1}},\overrightarrow{y_iy_{i-1}}) | N_{y_i} \cdot \overrightarrow{y_iy_{i+1}}|}{p(\overrightarrow{y_iy_{i+1}})} \right)
        \end{split}
    \end{equation}

The most effective way to use multiple importance samplingm in order to create a weight function is with the power heuristic:

\begin{equation}
	w_{s,t} = \frac{p_s^\beta}{\sum_{i=0}^{s+t-1} p_i^\beta}
\end{equation}

where the recommended value for $\beta = 2$

where $p_i$ is defined as the density for generating path $x_{s,t}$ using $i$ sub light path vertices and $s+t-i$ eye sub path vertices :

\begin{equation}
	p_i = p_{i,s+t-i}(\vec{x_{s,t}})
\end{equation}

The algorithm suggested in \cite{Vlnas2018} was implemented and is as follows:

\begin{algorithm}
\caption{BDPT}
	\begin{algorithmic}[1] % Remove the [1] if you do not want line numbers
		\Function{castRayBidirectional}{scene}
			define number of light and eye paths using Russian Roulette
		
		\ForAll{lightRays and cameraRays}
            \State $inter \gets intersect(ray)$
            \State $w\_, p \gets sample(inter)$
            \State $f = evaluateColor(inter, w\_)$
            \State $vertex.inter \gets inter$
            \State $vertex.beta \gets max(0, inter.normal \cdot w\_) \times f \times vertex_{previous}.beta / p / RR$
            \State $ray \gets ray(inter, w\_)$
		\EndFor
        \ForAll{Verticies}
            \State $p_{s, s+1} \gets \frac{p_{s-1, s} \times (w\__{s+1} \cdot w\__{s}) \times p_{vertex}} {(\|w\__s - w\__{s-1}\|)^2}$
        \EndFor
        \State $pSum \gets \sum_{s, t} (p_{camera\_s} \times p_{light\_t})^2$
        \State $hitColor \gets 0$
        \ForAll{Camera Verticies}
            \ForAll{Light Verticies}
                \If {Nothing blocking between verticies}
                    \State $w\_ \gets lightVertex.coords - cameraVertex.coords$
                    \State $\theta \gets max(0, inter.normal \cdot w\_)$
                    \State $connectingBeta \gets \frac{evaluateColor(cameraVertex, w\_) \times \theta}{p_{cameraVertex} \times RR}$
                    \State $G \gets \frac{|N_{cameraVertex} \cdot \overrightarrow{lightVertex.coordscameraVertex.coords}||N_{lightVertex.coords} \cdot \overrightarrow{cameraVertex.coordslightVertex.coords}|}{\|lightVertex.coords-cameraVertex.coords\|^2}$
                    \State $color \gets cameraVertex.beta \times lightVertex.beta \times connectingBeta \times G$
                    \State $hitColor \gets hitColor + \frac{color \times (p_{cameraVertex} \ times p_{lightVertex})^2}{pSum}$
                \EndIf
            \EndFor
        \EndFor
        \State return $hitColor$
        \EndFunction
	\end{algorithmic}
\end{algorithm}

\section*{Rendering Results and Experiments}


\section*{Discussion}


\bibliographystyle{plain}
\bibliography{references}

Generative AI statement: AI was used to help understand how MIS for bidirectional ray tracing worked and some minor debugging. No code was copied from AI sources

\end{document}
